pirouette: the error BEAST2 makes in inferring a phylogeny

### 2019-02-28 article.tex
Updates from Overleaf

Since BEAST performs a Bayesian analysis, 
it needs not only genetic data but also tree priors, 
describing how the diversification occur, to return posterior phylogenies.
BEAST works using standard tree priors like Yule or birth-death 
\giovanni{integrate this if necessary}.
The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
unlike its predecessor, the possibility for third-party users 
to integrate its prior libraries with new ones. 
Many of these have been developed, 
in the form of likelihood diversification models: 
birth-death models that have rates that are 
constant [\cite{nee1994reconstructed}], 
time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
diversity-dependent [\cite{etienne2011diversity}],
or can shift [the SLS model. Laudanno et al., in preparation].
There are birth death models in which speciation is a process that 
takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
can co-occur [Laudanno et al., in preparation],
dependent on a trait that is binary [\cite{maddison2007estimating}], 
has multiple states [\cite{fitzjohn2012diversitree}],
has a concealed state [\cite{beaulieu2016detecting}] 
or has multiple observed and concealed states [\cite{herrera2018detecting}].

\iffalse
Such models usually rely on the assumption 
that a particular diversification process 
can be explained mainly by a specific mechanism.
Such a mechanism is built into a likelihood function, 
which provides the probability that a phylogeny is realized 
from a given set of parameters.
Starting from a phylogeny, this likelihood is maximized
to infer the best estimates for the parameter values.
For such diversification models, it is a standard practice 
to measure how well parameters can be recovered in this way,
by simulating 'true' phylogenies from known parameters, then
measuring the parameters deduced by maximum likelihood. 
\fi

From a given phylogeny, the likelihood is maximized
to infer the best estimates for the parameter values.
For such  models, it is a standard practice to assess their performance 
on a controlled environment: this is done by simulating phylogenies from known parameters, 
then comparing the estimations obtained by maximum likelihood with the original parameters.
If the likelihood proves to be effective to describe the process, 
it can be implemented into BEAST2 as a new prior. 
Despite that it is still possible that already existing BEAST2 priors 
will perform well enough and, therefore, 
that the introduction of a new model is not necessary.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{workflow.png}
  \caption{
    \texttt{pirouette} pipeline. 
    The pipeline starts from a true phylogeny (1a), 
    that is converted to an alignment (2a), which is used
    to create a BEAST2 posterior of phylogenies (3a),
    which is compared to the true phylogeny (1a) using
    the nLTT statistic. (4a) shows the nLTT of the true tree in red,
    the nLTT of all individual posterior trees in grey, and the average
    nLTT statistic of all posterior trees in black.
    Taking the differences between the true tree's nLTT and each posterior's
    nLTT results in an error distribution, here displayed as a histogram
    of error values (5a).
    Optionally, a twin phylogeny (1b) can be generated from a true
    phylogeny (1a), which then results in an error distribution (5b) with
    the similar intermediate steps 
  }
  \label{fig:pipeline}
\end{figure}

@@

The first step simulates a DNA alignment from the given 
phylogeny (Fig.~\ref{fig:pipeline} 1a $\rightarrow$ 2a).
One can specify a DNA sequence
of any length at the root of the phylogeny, a DNA mutation rate, a
site (i.e. nucleotide substitution) model, 
a clock model, a random number generator (RNG) seed and a location
where the alignment is saved to. This step is relatively fast, but longer
DNA alignments will noticeably slow down the inference step.

The second step selects an inference model, using the alignment.
The user can specify whether to use the generative model or select the best model
from a set of inference models. 
When selecting the generative model,
the site and clock model used in the alignment simulation are used
in the inference. Because the given phylogeny (on which the alignment is based)
may have followed any tree prior (i.e. speciation model), the user needs
to specify which tree prior is used in the inference. 
When selecting for the best
model, the alignment is used to find the inference model that has the
highest evidence (i.e. marginal likelihood) from a set of candidate inference models.
The evidence of an inference model is estimated using a nested sampling (NS)
approach, as described in \cite{maturana2018model}. The nested sampling is
performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
inference model selection is not available on Windows.

The third step infers a Bayesian posterior from the simulated 
alignment (Fig.~\ref{fig:pipeline} 2a $\rightarrow$ 3a),
using the inference model(s) selected in the previous step. The user
can specify the additional parameters needed for the BEAST2 run, which
are the Markov-Chain Monte Carlo (MCMC) setup, 
an optional MRCA prior and an RNG seed.
The MCMC setup determines the number of posterior trees sampled.
An MRCA prior allows the inferred phylogeny to have a dated crown age.

The fourth step compares the true tree to each posterior tree
error using
the nLTT statistic (\cite{janzen2015approximate}) 
(Fig.~\ref{fig:pipeline} 3a $\rightarrow$ 4a), resulting
in an error distribution (Fig.~\ref{fig:pipeline} 4a $\rightarrow$ 5a).
\richel{I would not mind to remove 4a and 4b from the figure}
The nLTT statistic is used by default,
but also a user-defined error statistic can be used. As an example,
\verb;pirouette; suplies such a custom error statistic,
that uses an absolute difference in the gamma statistic [\cite{pybus2000testing}].
Additionally, the user can specify the
proportion of posterior phylogenies to 
discard (i.e. the burn-in), throwing away the first $10\%$
of all phylogenies by default. This burn-in is used to discard
the part of the MCMC chain that has not yet converged to a
representative part of the state space.

@@

An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline} 1a $\rightarrow$ 1b), 
that will be analyzed in the same way as the true tree.
The twin tree has the same topology as the given phylogeny, 
yet its branching times are simulated according to 
the tree prior used by BEAST2 to produce a posterior.
The goal of this procedure is to generate a tree 
to use as a control for the main experiment.
The procedure to generate a twin tree is the following:

@@

An overview of \verb;pirouette;'s main functions is shown in 
table \ref{tab:functions}, which are all demonstrated in the examples
below.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: first research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A first research question that \verb;pirouette; answers is:
"What is the error made by BEAST2 from a phylogeny using the same diversification model as it was generated by?"

@@

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_1.png}
  \caption{
    Example 1: the error BEAST2 makes from a phylogeny 
    when the generative and inference model are the same.
    The x-axis specifies the tree type (which can be 'true' or 'twin'),
    the y-axis shows the error density
  }
  \label{fig:example_1}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: second research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A second research question that \verb;pirouette; answers, is:
"What is the error BEAST2 makes from a phylogeny when
picking the best inference model?"

@@

In the second step we state our experiments. 
Within this example, all our experiments are candidates,
run only if it is the best candidate, the evidence is measured (ignoring
this will give a helpful error) and we use the full arsenal of 
40 inference models, which are all combinations of 4 site 
models, 2 clock models and 5 tree priors.

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_2.png}
  \caption{
    Example 2: the error BEAST2 makes from a phylogeny when
    picking the best inference model.
    The x-axis specifies the tree type (which can be 'true' or 'twin'),
    the y-axis shows the error density
  }
  \label{fig:example_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: third research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A third research question that \verb;pirouette; answers is:
"What is the error BEAST2 makes from a phylogeny, 
when hand-picking an inference model, compared to the background noise?"

The following settings are the same as in the previous section:
phylogeny (listing \ref{lst:unknown_phylogeny}), 
alignment parameters (listing \ref{lst:create_alignment}), 
experiments (listing \ref{lst:create_generative_experiment}),
BEAST2 inference and error measuring parameters.

\iffalse
This time, we are interested in creating a twin tree. A twin tree
has the same topology as the given tree, yet its branch lengths follow
a distribution as such that the BEAST2 error will be 
close to \richel{'close to' is a bit weasly, but currently it is true}
its minimal error.
\fi

This time, we are interested in creating a twin tree. A twin tree
has the same topology as the given tree, 
yet its branching times are simulated according 
to the tree prior used by BEAST2 to produce a posterior. 
Since there is coherence between the inference model and the generative model, 
we expect BEAST2 to produce a posterior of phylogenies clearly 
more similar to the original one. 
We intend to use this as a control experiment.

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_3.png}
  \caption{
    Example 3: the error BEAST2 makes from a phylogeny picking the best inference model versus the control.
    Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
    The x-axis specifies the tree type (which can be 'true' or 'twin'),
    the y-axis shows the error density
  }
  \label{fig:example_3}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: fourth research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A fourth research question that \verb;pirouette; answers, is:
"What is the error BEAST2 makes from a phylogeny for a known
generative model, compared to candidates with similar inference models,
in relationship to the background error?"

Here, same as example 2, we start with a tree generated by the Yule process,
by using the code shown in listing \ref{lst:create_yule_tree}.
The first step in \verb;pirouette; is to simulate a DNA alignment 
from the given phylogeny, for which we'll use the same code 
as shown in listing \ref{lst:create_alignment}.

@@

In the second step we state our experiments. 
Within this example, we must both specify the generative and candidate 
experiments. Here we'll specify the generative model:

@@

Specifying all candidates can be done in a factorial fashion using 
\verb;create_all_experiments;. In this case, all candidates will use the
same site and clock model, and only differ in their tree priors:

@@

Also here, the third (the BEAST2 inference) and fourth (measuring the error)
steps have sensible defaults. To get an idea of a baseline error made by
BEAST2, we also make use of the twinning option:

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_4.png}
  \caption{
    Example 4: the error BEAST2 makes from a phylogeny for a known
    generative model, compared to candidates with similar inference models,
    in relationship to the background error.
    The x-axis specifies the tree type,
    the y-axis shows the error density
  }
  \label{fig:example_4}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | r | l | l | l | l | }
\hline
\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
\cline{2-5}
                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
\hline
1  & $0.02$  & $0.0 - 0.27$  &         &               \\
2  & $0.02$  & $0.0 - 0.08$  &         &               \\
3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
\hline
\end{tabular}
\caption{Results of the examples. 
  For the true and twin tree,
  the median and range of the errors 
  is shown. Experiment 4 is split in two rows
  by either the candidate ('C') or generative model ('G') 
}
\label{tab:results}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{check numbers}

Here we discuss how to interpret the results of \verb;pirouette;.
Figure \ref{fig:example_1} shows that the generative model 
gives an error distribution with a median of \richel{?}
The range of errors goes from \richel{?} to \richel{?}.
As we know the generative model, 
this error distribution is the minimum noise that BEAST2 produces 
for that generative model.
We can measure the effective samples as well (not shown), 
and all ESSes are above
the recommended value of 200 \cite{drummond2015bayesian}, 
as shown in table \ref{tab:esses_example_1}.

Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution 
with a median of \richel{?}. 
The range of errors goes from \richel{?} to \richel{?}. 
The inference model with the highest evidence follows a JC site model, 
RLN clock model and BD tree prior, with a weight of $0.675$. 
This is (unexpectedly) better than the generative
model with a weight of $0.166$. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_2}.
This means that it is rational to use this inference model to 
infer a posterior phylogeny distribution, 
as is done by \verb;pirouette;.
Also here, the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_2}.

@@

Figure \ref{fig:example_4} shows that \richel{?}
and 
gives an error distribution with a median of \richel{?}, 
The twin tree has an error distribution with a median of \richel{?},
with a range from 0.0 to \richel{?}, which is an indication of the
error made by BEAST on that tree prior.
The inference model with the highest evidence follows a JC site model, 
RLN clock model and BD tree prior, with a weight of ?. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_4}.
As there is little difference between the true and twin tree,
the (unknown) diversification model has little impact on the error
BEAST2 makes, hinting that the constant-rate birth-death model is
sufficiently good enough to explain the true phylogeny. 
The diversification model apparently has minor impact and need not to be
added to BEAST2.
We can draw this conclusion with confidence, as the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_4}.

>>>>

Since BEAST performs a Bayesian analysis, 
it needs not only genetic data but also tree priors, 
describing how the diversification occur, to return posterior phylogenies.
BEAST works using standard tree priors like Yule or birth-death 
\giovanni{integrate this if necessary}.
The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
unlike its predecessor, the possibility for third-party users 
to integrate its prior libraries with new ones. 
Many of these have been developed, 
in the form of likelihood diversification models \giovanni{Not of them are already available for BEAST2 yet, or maybe they will never be}: 
birth-death models that have rates that are 
constant [\cite{nee1994reconstructed}], 
time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
diversity-dependent [\cite{etienne2011diversity}],
or can shift [\cite{etienne2012conceptual}, the SLS model. Laudanno et al., in preparation].
There are birth death models in which speciation is a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], can co-occur [Laudanno et al., in preparation],
dependent on a trait that is binary [\cite{maddison2007estimating}], 
has multiple states [\cite{fitzjohn2012diversitree}],
has a concealed state [\cite{beaulieu2016detecting}] 
or has multiple observed and concealed states [\cite{herrera2018detecting}].

From a given phylogeny, the likelihood is maximized
to infer the best estimates for the parameter values.
For such  models, it is a standard practice to assess their performance 
on a controlled environment: this is done by simulating phylogenies from known parameters, 
then comparing the estimations obtained by maximum likelihood with the original parameters.
If the likelihood proves to be effective to describe the process, 
it can potentially be implemented into BEAST2 as a new prior. 
Despite that, it is still possible that already existing BEAST2 priors 
will perform well enough and, therefore, 
that the introduction of a new model is not necessary.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{workflow.png}
  \caption{
    \texttt{pirouette} pipeline. 
    The pipeline starts from a true phylogeny (1a), 
    that is converted to an alignment (2a), which is used
    to create a BEAST2 posterior of phylogenies (3a),
    which is compared to the true phylogeny (1a) \giovanni{this makes me think that we might need an arrow going from 1a to 4a, maybe (and i say maybe) the arrow could be in red to highlight that the "true" nltt is in red too. BTW, the red line is not supervisible.} using
    the nLTT statistic. (4a) shows the nLTT of the true tree in red,
    the nLTT of all individual posterior trees in grey, and the average
    nLTT statistic of all posterior trees in black.
    Taking the differences between the true tree's nLTT and each posterior's
    nLTT results in an error distribution, here displayed as a histogram
    of error values (5a).
    Optionally, a twin phylogeny (1b) can be generated from a true
    phylogeny (1a), which then results in an error distribution (5b) with
    the similar intermediate steps 
  }
  \label{fig:pipeline}
\end{figure}

@@

The first step simulates a DNA alignment from the given 
phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
One can specify a DNA sequence
of any length at the root of the phylogeny, a DNA mutation rate, a
site (i.e. nucleotide substitution) model, 
a clock model, a random number generator (RNG) seed and a location
where the alignment is saved to. This step is relatively fast, but longer
DNA alignments will noticeably slow down the inference step.

The second step selects an inference model, using the alignment \giovanni{Is it worth to add it in the figure, maybe on the side of the vertical pipeline? It could be something like an arrow moving horizontally to the box of inference model. Then the box could have another arrow pointing at the arrow connecting 2 and 3. If you think this is too complicated, and not easy to follow for the reader you can discard it.}.
The user can specify whether to use the generative model or select the best model from a set of inference models. 
When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Because the given phylogeny (on which the alignment is based)
may have followed any tree prior (i.e. speciation model), the user needs
to specify which tree prior is used in the inference. 
When selecting for the best
model, the alignment is used to find the inference model that has the
highest evidence (i.e. marginal likelihood) from a set of candidate inference models.
The evidence of an inference model is estimated using a nested sampling (NS)
approach, as described in \cite{maturana2018model}. The nested sampling is
performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
inference model selection is not available on Windows.

The third step infers a Bayesian posterior from the simulated 
alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a),
using the inference model(s) selected in the previous step. The user
can specify the additional parameters needed for the BEAST2 run, which
are the Markov-Chain Monte Carlo (MCMC) setup, 
an optional MRCA prior and an RNG seed \giovanni{This has been already said when defining the inference model. I suggest to define it only once in detail (like here) and just mention it the first time (something like: "An inference model is a combination of site model, clock model, tree prior as well as additional BEAST2 settings." in the description section, around line 98), which will probably look a bit more neat, as it's a bit clearer the parallel between alignment model and inference model.}.
The MCMC setup determines the number of posterior trees sampled.
An MRCA prior allows the inferred phylogeny to have a dated crown age.

The fourth step compares the true tree to each posterior tree
error using the nLTT statistic (\cite{janzen2015approximate}) 
(Fig.~\ref{fig:pipeline}, 3a $\rightarrow$ 4a), resulting
in an error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
\richel{I would not mind to remove 4a and 4b from the figure}
The nLTT statistic is used by default, but also a user-defined error statistic can be used. As an example,
\verb;pirouette; supplies one custom error statistic,
that uses an absolute difference in the gamma statistic [\cite{pybus2000testing}].
Additionally, the user can specify the
proportion of posterior phylogenies to 
discard (i.e. the burn-in), throwing away the first $10\%$
of all phylogenies by default. This burn-in is used to discard
the part of the MCMC chain that has not yet converged to a
representative part of the state space.

@@

An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
that will be analyzed in the same way as the true tree.
The twin tree has the same topology as the given phylogeny, 
yet its branching times are simulated according to 
the tree prior used by BEAST2 to produce a posterior \giovanni{is it true? is it Yule or BD? Please check}.
The goal of this procedure is to generate a tree 
to use as a control for the main experiment.
We produce the twin tree according to the following procedure.

@@

An overview of \verb;pirouette;'s main functions is shown in 
table \ref{tab:functions}. We will exploit these functions to answer our research questions in the next sections.

@@

All \verb;pirouette;'s functions are documented. Useful examples and sensible defaults, which we will refer to in the next sections, can be accessed from the documentation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: First research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A first research question that \verb;pirouette; answers is:

"What is the error made by BEAST2 from a phylogeny using the same diversification model as it was generated by?"

@@

\giovanni{do we really want to use a function called "create\_ideal\_tree"?}

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_1.png}
  \caption{
    Example 1: the error BEAST2 makes from a phylogeny 
    when the generative and inference model are the same.
    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error. \giovanni{This holds true for all the figures: if we want to add these numbers we need to specify what they represent in the caption.}
  }
  \label{fig:example_1}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Second research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A second research question that \verb;pirouette; answers, is:

"What is the error made by BEAST2 from a phylogeny when
picking the best inference model?"

@@

In the second step we state our experiments. 
Within this example, all our experiments are candidates,
run only if it is the best candidate, the evidence is measured (ignoring
this will give a helpful error) and we use the full set of 
40 inference models, which are all combinations of 4 site 
models, 2 clock models and 5 tree priors.

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_2.png}
  \caption{
    Example 2: the error BEAST2 makes from a phylogeny when
    picking the best inference model.
    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
  }
  \label{fig:example_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Third research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A third research question that \verb;pirouette; answers is:

"What is the error made by BEAST2 from a phylogeny, 
when hand-picking an inference model, compared to the background noise?"

The following settings are the same as in the previous section:
phylogeny (listing \ref{lst:unknown_phylogeny}), 
alignment parameters (listing \ref{lst:create_alignment}), 
experiments (listing \ref{lst:create_generative_experiment}),
BEAST2 inference and error measuring parameters \giovanni{can we add a reference to the "BEAST2 inference and error measuring parameters" we are using?}.

This time, we are interested in creating a twin tree. A twin tree
has the same topology as the given tree, 
yet its branching times are simulated according 
to the tree prior used by BEAST2 to produce a posterior. 
Since there is coherence between the inference model and the generative model \giovanni{Do we need to specify what is a generative model? Are we using BD or Yule? In case we are using Yule are we sure there is actually coherence when inference model is bd?}, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one. 
We intend to use this as a control experiment.

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_3.png}
  \caption{
    Example 3: the error made by BEAST2 from a phylogeny, picking the best inference model versus the control.
    Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
  }
  \label{fig:example_3}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Fourth research question}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A fourth research question that \verb;pirouette; answers, is:

"What is the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error?"
\giovanni{Did we define already the "generative" model? If not we can maybe refer to it when we define alignment and inference model}

Here, same as example 2, we start with a tree generated by the Yule process, by using the code shown in listing \ref{lst:create_yule_tree}.
The first step in \verb;pirouette; is to simulate a DNA alignment 
from the given phylogeny, for which we will use the same code 
as shown in listing \ref{lst:create_alignment}.

@@

In the second step we state our experiments. 
In this example, we must both specify the generative and candidate experiments. Here we specify the generative model:

@@

Specifying all candidates can be done simply by calling \verb;create_all_experiments;. In this case, all candidates will use the same site and clock model, and only differ in their tree priors:

@@

We choose again to use sensible defaults for the third (the BEAST2 inference) and fourth (measuring the error) steps, as we previously did to answer the second research question. To get an idea of a baseline error made by BEAST2, we also make use of the twinning option:

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_example_4.png}
  \caption{
    Example 4: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
    The x-axis specifies the tree type, the y-axis shows the error.
  }
  \label{fig:example_4}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | r | l | l | l | l | }
\hline
\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
\cline{2-5}
                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
\hline
1  & $0.02$  & $0.0 - 0.27$  &         &               \\
2  & $0.02$  & $0.0 - 0.08$  &         &               \\
3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
\hline
\end{tabular}
\caption{Results of the examples. 
  For the true and twin tree,
  the median and range of the errors 
  is shown. Experiment 4 is split in two rows
  by either the candidate ('C') or generative model ('G'). \giovanni{How do you define range? Are we using the standard error? \url{https://en.wikipedia.org/wiki/Standard_error}}
}
\label{tab:results}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{check numbers}
\giovanni{I think that the current version of the discussion is too didascalic. Most of what is said can be effectively summarized by table 5. We can even add ESSes in it. Another point is that we should understand what is the message we want to deliver in this discussion. Since we are only using one tree we are not in the position to claim anything, because a single tree is not an effective sample. I saw that in some occasions you refer to the RQ as examples. This might be the right way to proceed, as our purpose is just to illustrate how to run all the possible experiments on a full tree distribution. I think we should make it clearer throughout the entire manuscript. I am also interested in hearing Rampal's opinion on this.}

Here we discuss how to interpret the results of \verb;pirouette;.
Figure \ref{fig:example_1} shows that the generative model 
gives an error distribution with a median of \richel{?}
The range of errors goes from \richel{?} to \richel{?}.
As we know the generative model, this error distribution is the minimum noise that BEAST2 produces for that generative model.
We can measure the effective samples as well (not shown), 
and all the Estimated Sample Sizes (ESSes) are above
the recommended value of 200 [\cite{drummond2015bayesian}], 
as shown in table \ref{tab:esses_example_1}.

Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution with a median of \richel{?}. 
The range of errors goes from \richel{?} to \richel{?}. 
The inference model with the highest evidence follows a JC site model, RLN clock model and BD tree prior, with a weight of $0.675$. 
This is (unexpectedly) \giovanni{We are not using distributions, so we are totally subject to stochasticity. Nothing can be expected from just one tree.} better than the generative
model with a weight of $0.166$. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_2}.
This means that it is rational to use this inference model to 
infer a posterior phylogeny distribution, as is done by \verb;pirouette;.
Also here, the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_2}.

@@

Figure \ref{fig:example_4} shows that \richel{?}
and 
gives an error distribution with a median of \richel{?}, 
The twin tree has an error distribution with a median of \richel{?},
with a range from 0.0 to \richel{?}, which is an indication of the
error made by BEAST on that tree prior.
The inference model with the highest evidence follows a JC site model, 
RLN clock model and BD tree prior, with a weight of ?. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_4}.
As there is little difference between the true and twin tree,
the (unknown) diversification model has little impact on the error
BEAST2 makes, hinting that the constant-rate birth-death model is
sufficiently good enough to explain the true phylogeny. 
The diversification model apparently has minor impact and need not to be
added to BEAST2.
We can draw this conclusion with confidence \giovanni{no, we cannot: one tree is not enough}, as the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_4}.
